Correlation between photons emitted in different channels from an atom inside a leaky driven cavity

### 2021-08-11 EPJD/newpaper.tex
Modificacion misteriosa

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps, 
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=blue,
}
\usepackage[ruled,vlined]{algorithm2e}

\usepackage{physics}
\usepackage[draft,inline,nomargin]{fixme} \fxsetup{theme=color}
%\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

%\usepackage[showframe,%Uncomment any one of the following lines to test 
%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
%%text={7in,10in},centering,
%%margin=1.5in,
%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
%%height=10in,a5paper,hmargin={3cm,0.8in},
%]{geometry}

\begin{document}

\preprint{APS/123-QED}

\title{Correlation between photons emitted in diferent channels from
  an atom inside a leaky driven cavity}% Force line breaks with \\

\author{Guillermo Preisser}
 \email{Corresponding author e-mail: guillermo.preisser@uabc.edu.mx}
 %Lines break automatically or can be forced with \\
\author{Pablo Barberis-Blostein}%
\affiliation{Instituto de Investigaciones en Matemáticas Aplicadas y en Sistemas, Universidad Nacional Autónoma de México, Ciudad Universitaria, 04510, CDMX, México.}

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
  Photons emitted by a driven two level atom in free space do not show
  correlation between different directions. We consider a two level
  atom inside a driven leaky cavity. We consider two atomic emission
  channels: emission to cavity modes and emission to no-cavity modes.
  Using the quantum trajectory formalism we study the correlation
  between photons emitted into cavity modes and that leak the cavity
  through its mirror, with photons that are emitted into the other
  modes. We find that the cavity generates a correlation between the
  two emission channels.
\end{abstract}

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

@@

Spontaneous emission has been studied since the early XX century
\cite{10.2307/94746, 1917PhyZ...18..121E}. One of its main
characteristics is that the exact time of emission, as well as the
direction in which the photon is emitted, is random, which difficults
its measurement. On the other hand, photons that leak a cavity mirror
are easier to measure because we know their direction \cite{326305,
  doi:10.1063/1.113345}. The randomness of the direction in which the
photons are emitted by atoms in free space difficults photon counting
experiments. To sort out this difficulty one could think of a physical
arrangement in which a two-level atom is coupled to a driven cavity,
as it is shown in Fig. \ref{asa}. In this physical arrangement, we
have two emission channels: spontaneous emissions of the atom
--photons that are emitted in modes that are not cavity modes and
therefore escape the cavity to its side-- and cavity emissions
--photons that leak through one of the cavity mirrors. In the case
that a correlation exists between the two channels, we can count the
number of photons that leak the cavity through its mirror, which is
easy to measure, to find information about the number of photons
emitted into the other modes (spontaneous emission), which are
difficult to measure.

@@

We use quantum trajectories to calculate the joint probability of the
number of spontaneus emissions and the number of cavity emissions. A
quantum trajectory describes a continuously monitored quantum system
\cite{Carmichael1993Open}. In our case each trajectory gives the
number of photons emitted in each channel. Simulating a large number
of trajectories and registering the number of photon emissions into
the two channels, we obtain its joint probability distribution and
their correlation. The correlation varies with the cavity linewidth
and the atom-cavity coupling. For a very bad cavity we recover the
free space result: there is no correlation between the two channels.

@@

\section{Theoretical model}\label{sc:drivenjc}
We consider a system consisting of a two-level atom, coupled to a mode
of a leaky cavity. The mode is driven by a coherent field resonant
with the atomic transition and the cavity mode.
% In order to account for
% dissipation one may use a non-hermitian Hamiltonian to describe the
% system \cite{bla}. Using this description for the system will turn
% useful at the moment of using quantum trajectory theory. Such
% Hamiltonian can be separated into hermitian and non-hermitian terms.
%\begin{equation}
%H = H_s - \frac{i\hbar}{2}\sum_m C^\dagger_n C_n .
%\end{equation}
The Hamiltonian of this system, in the rotating wave approximation, is
given by
\begin{equation}
H_s = H_{at} + H_{cav} + H_{atcav} + H_{lascav}, \label{mainham}  
\end{equation}
where the atom and cavity terms are, respectively
\begin{subequations}
\begin{equation}
H_{at} = \tfrac{1}{2}\hbar \omega \sigma_z,    
\end{equation}
\begin{equation}
H_{cav} = \hbar \omega  a^\dagger a,  
\end{equation}
$\omega$ is the cavity mode frequency and the atomic transition
frequency. The atom is in resonance with the cavity mode. 
The coupling between atom and cavity, and cavity and coherent
field are
\begin{equation}
H_{atcav} = i\hbar g(a\sigma_+ - a^\dagger \sigma_-),    
\end{equation}
\begin{equation}
H_{lascav} = i\hbar \mathcal{E}(ae^{i\omega t} - a^\dagger e^{-i\omega t}),    
\end{equation}
\end{subequations}
where the annihilation operator $a$ of a single photon in the cavity,
and the creation operator $a^\dagger$ obey $[a, a^\dagger] = 1$. The
operators $\sigma_-, \sigma_+$ account for the atomic lowering and
raising operators, respectively, and the operator $\sigma_z$ accounts
for the population difference between the atomic levels. These
operators obey
$[\sigma_+, \sigma_-] = \sigma_z, \ [\sigma_\pm, \sigma_z] = \mp
2\sigma_\pm$. The strength of the coupling between atom and cavity
field is given by the constant
\begin{equation}
g = \sqrt{\frac{\omega d^2}{2\hbar \epsilon_0 V_Q}},
\end{equation}
where $V_Q$ is the mode volume, $d$ is the atomic dipole, and
$\epsilon_0$ is the permittivity of free space. $\mathcal{E}$ is a
term proportional to the coherent field that drives the cavity in
resonance. Dissipation is introduced by the following non-hermitian
Hamiltonian
\begin{equation} \label{nh}
H_{nh} = H_{dec} + H_{leak},
\end{equation}
where the terms corresponding to spontaneous emission of the atom and cavity losses are
\begin{subequations}
\begin{equation}
H_{dec} = - i\hbar\frac{\gamma}{2}\sigma_+\sigma_-,
\end{equation}
\begin{equation}
H_{leak} = - i\hbar\kappa a^\dagger a\, ,
\end{equation}
\end{subequations}
where $\kappa$ and $\gamma$ are the cavity loss and spontaneous decay
rate. The evolution between jumps is given by $H=H_s+H_{nh}$. This
evolution is interrupted by spontaneous emission by the atom or when a
photon leaks the cavity.

@@

\begin{subequations}\label{relprob}
  \begin{eqnarray}  \delta p_1 &=& \gamma \bra{\phi
      (t)} \sigma_+
                                                  \sigma_-\ket{\psi (t)}\delta t\, , \\
    \delta p_2 &=& 2\kappa \bra{\phi (t)} a^\dagger a\ket{\psi
                   (t)}\delta t \, .
  \end{eqnarray}
\end{subequations}
\makeatletter
%\DeclareCaptionLabelFormat{numberless}{\ALG@name#1}
%\captionsetup[algorithm]{labelformat=numberless} 
\makeatother
\begin{algorithm}
\caption{Pseudocode for register of cavity losses and atom's spontaneous emissions}
\label{alg1}
\begin{algorithmic}
%\caption{Pseudocode for register of cavity losses and atom's spontaneous emissions}
\STATE $N$: number of temporal intervals
\STATE $N_A $: number of spontaneous emissions
\STATE $N_C$: number of cavity emissions
\STATE $N_{max}$: maximum number of cavity emissions
\STATE $N_{CA}$: matrix relating the number of cavity emissions with the number of spontaneous emissions

@@

\ENDIF
\ELSE
\STATE $\ket{ \phi(t + \delta t)} = \frac{\ket{\phi^1 (t+ \delta t)}}{\sqrt{1-\delta p}}$
\STATE $t \ += \delta t$
\ENDIF
\ENDFOR
\RETURN  $N_{CA}$
\end{algorithmic}
\end{algorithm}
We compare the collapse probability of Eq.~\eqref{probcol} with a
random number $\epsilon$, uniformly distributed between zero and one.
In the case the condition $\delta p > \epsilon$ is fulfilled we
simulate that a quantum jump has happened, by acting with one of the
collapse operators $a$ or $\sigma_-$ and renormalizating the obtained
state. The operator that will act on the state will be given according
to the relative probabilities of Eqs.~\eqref{relprob}.
\begin{figure}[!t] 
\centering
\includegraphics[scale = 0.5]{distributioneng.pdf}
\caption{Probability distribution of spontaneous emissions for
  different number of cavity photons. Parameters: $\expval{n} \approx 1$, $\kappa = g = 0.1\gamma$. } \label{probdiss}
\end{figure}
If $\delta p > \epsilon$ is not fulfilled we just renormalize the
state, because using a non-hermitian Hamiltonian will have as a
consequence that the probability won't be conserved
\cite{Sakurai:1167961}. For the next step we repeat the procedure
using the new state. One quantum trajectory of total time $t_f$ is
obtained after repeating the previous procedure $t_f/\delta t$ steps.
For each trajectory we count the number of quantum jumps, the jumps
corresponding to operator $a$ reduces the number of photons inside the
cavity and corresponds to a photon that leak the cavity. The jumps
corresponding to operator $\sigma_-$, send the atom back to the ground
state and corresponds to a spontaneous emission. The procedure to
register emissions in the two different channels is summarized in
algorithm. \ref{alg1}.

@@

\begin{center}
\begin{figure}[h!]
\begin{center}
\includegraphics[scale = 0.5]{million2.pdf}
\caption{Pearson correlation coefficient as a function of $\kappa$ for
  quantum trajectories with different duration. Parameters:
  $\expval{n} \approx 1$, $g = 0.1\gamma$.} \label{errorzz}
\end{center}
\end{figure}
\end{center}
%kncrease or increases

%\begin{center}
%\begin{figure}[h!]
%\begin{center}
%\includegraphics[scale = 0.7]{lastgraph.pdf}
%\caption{\small{Mean value of spontaneous emissions after post-selection, fixing $\expval{n} \approx 1$, $\gamma =$ 1.0, $g = \kappa =$ 0.1, and with  $\mathcal{E} =  \kappa |\langle n \rangle|[1 + 2g^2/\gamma \kappa]$.}}  \label{errorzz}
%\end{center}
%\end{figure}
%\end{center}

@@

% In this work we considered a driven Jaynes-Cummings system. We made a
% program in which, through the use of quantum trajectory theory, we
% were able to obtain probability distributions that relate the number
% of spontaneous emissions given certain number of cavity emissions.
% Through the probability distributions we obtained estimations for the
% number of spontaneous emissions by measuring cavity emissions,
% considering the correspondent error. Finally we study the existing
% correlation where we saw how this correlation decreases as we increase
% the cavity linewidth. However, if we try with values for
% $\kappa/\gamma < 0.1$, there comes a point where the increase stops
% and we see a decrease. We believe this is related to the fact that if
% we use enough small values of $\kappa$ there will come a point where
% the emission of photons will be too low and this will imply a decrease
% in the correlation. The results in this work provide a tool that could
% be useful in the manipulation of single photons product of spontaneous
% emission, and provide insight into the physical behavior of the driven
% Jaynes-Cummings model.

\begin{acknowledgments}
This work was supported by DGAPA-UNAM under grant PAPIIT-IG120518.
\end{acknowledgments}

\bibliography{reftesis}% Produces the bibliography via BibTeX.

>>>>

\documentclass[epj,final]{svjour}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=blue,
}
\usepackage[ruled,vlined]{algorithm2e}

\usepackage{physics}
\usepackage{algorithmic}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage[square,numbers]{natbib}
\bibliographystyle{unsrtnat}
\begin{document}

\title{Correlation between photons emitted in different channels from
  an atom inside a leaky driven cavity}% Force line breaks with \\

\author{Guillermo Preisser \and  Pablo Barberis-Blostein}
 \mail{guillermo.preisser@uabc.edu.mx}

%\offprints{}
 \institute{Instituto de Investigaciones en Matem\'aticas Aplicadas y en
   Sistemas, Universidad Nacional Aut\'onoma de M\'exico, Ciudad
   Universitaria, 04510, CDMX, México.}

\date{\today}

\abstract{
  Photons emitted by a driven two level atom in free space do not show
  correlation between different directions. We consider a two level
  atom inside a driven leaky cavity. We consider two atomic emission
  channels: emission to cavity modes and emission to no-cavity modes.
  Using the quantum trajectory formalism we study the correlation
  between photons emitted into cavity modes and that leak the cavity
  through its mirror, with photons that are emitted into the other
  modes. We find that the cavity generates a correlation between the
  two emission channels.}

\titlerunning{Correlations in an atom inside a leaky driven cavity}
\maketitle

@@

Spontaneous emission has been studied since the early XX century
\cite{10.2307/94746} \cite{1917PhyZ...18..121E}. One of its main
characteristics is that the exact time of emission, as well as the
direction in which the photon is emitted, is random, which difficults
its measurement. On the other hand, photons that leak a cavity mirror
are easier to measure because we know their direction
\cite{326305} \cite{doi:10.1063/1.113345}. The randomness of the direction in
which the photons are emitted by atoms in free space difficults photon
counting experiments. To sort out this difficulty one could think of a
physical arrangement in which a two-level atom is coupled to a driven
cavity, as it is shown in Fig. \ref{asa}. In this physical
arrangement, we have two emission channels: spontaneous emissions of
the atom --photons that are emitted in modes that are not cavity modes
and therefore escape the cavity to its side-- and cavity emissions
--photons that leak through one of the cavity mirrors. In the case
that a correlation exists between the two channels, we can count the
number of photons that leak the cavity through its mirror, which is
easy to measure, to find information about the number of photons
emitted into the other modes (spontaneous emission), which are
difficult to measure.

@@

We use quantum trajectories to calculate the joint probability of the
number of spontaneous emissions and the number of cavity emissions. A
quantum trajectory describes a continuously monitored quantum system
\cite{Carmichael1993Open}. In our case each trajectory gives the
number of photons emitted in each channel. Simulating a large number
of trajectories and registering the number of photon emissions into
the two channels, we obtain its joint probability distribution and
their correlation. The correlation varies with the cavity linewidth
and the atom-cavity coupling. For a very bad cavity we recover the
free space result: there is no correlation between the two channels.

@@

\section{Theoretical model}\label{sc:drivenjc}
We consider a system consisting of a two-level atom, coupled to a mode
of a leaky cavity. The mode is driven by a coherent field resonant
with the atomic transition and the cavity mode.
The Hamiltonian of this system, in the rotating wave approximation, is
given by
\begin{equation}
H_s = H_{at} + H_{cav} + H_{atcav} + H_{lascav}, \label{mainham}  
\end{equation}
where the atom and cavity terms are, respectively
\begin{subequations}
\begin{equation}
H_{at} = \tfrac{1}{2}\hbar \omega \sigma_z,    
\end{equation}
\begin{equation}
H_{cav} = \hbar \omega  a^\dagger a,  
\end{equation}
$\omega$ is the cavity mode frequency and the atomic transition
frequency. The atom is in resonance with the cavity mode. 
The coupling between atom and cavity, and cavity and coherent
field are
\begin{equation}
H_{atcav} = i\hbar g(a\sigma_+ - a^\dagger \sigma_-),    
\end{equation}
\begin{equation}
H_{lascav} = i\hbar \mathcal{E}(ae^{i\omega t} - a^\dagger e^{-i\omega t}),    
\end{equation}
\end{subequations}
where the annihilation operator $a$ of a single photon in the cavity,
and the creation operator $a^\dagger$ obey $[a, a^\dagger] = 1$. The
operators $\sigma_-, \sigma_+$ account for the atomic lowering and
raising operators, respectively, and the operator $\sigma_z$ accounts
for the population difference between the atomic levels. These
operators obey
$[\sigma_+, \sigma_-] = \sigma_z, \ [\sigma_\pm, \sigma_z] = \mp
2\sigma_\pm$. The strength of the coupling between atom and cavity
field is given by the constant
\begin{equation}
g = \sqrt{\frac{\omega d^2}{2\hbar \epsilon_0 V_Q}},
\end{equation}
where $V_Q$ is the mode volume, $d$ is the atomic dipole, and
$\epsilon_0$ is the permittivity of free space. $\mathcal{E}$ is a
term proportional to the coherent field that drives the cavity in
resonance. Dissipation is introduced by the following non-hermitian
Hamiltonian
\begin{equation} \label{nh}
H_{nh} = H_{dec} + H_{leak},
\end{equation}
where the terms corresponding to spontaneous emission of the atom and cavity losses are
\begin{subequations}
\begin{equation}
H_{dec} = - i\hbar\frac{\gamma}{2}\sigma_+\sigma_-,
\end{equation}
\begin{equation}
H_{leak} = - i\hbar\kappa a^\dagger a\, ,
\end{equation}
\end{subequations}
where $\kappa$ and $\gamma$ are the cavity loss and spontaneous decay
rate. The evolution between jumps is given by $H=H_s+H_{nh}$. This
evolution is interrupted by spontaneous emission by the atom or when a
photon leaks the cavity.

@@

\begin{subequations}\label{relprob}
  \begin{eqnarray}  \delta p_1 &=& \gamma \bra{\phi
      (t)} \sigma_+
                                                  \sigma_-\ket{\psi (t)}\delta t\, , \\
    \delta p_2 &=& 2\kappa \bra{\phi (t)} a^\dagger a\ket{\psi
                   (t)}\delta t \, .
  \end{eqnarray}
\end{subequations}
\makeatletter
\makeatother
\begin{algorithm}
\caption{Pseudocode for register of cavity losses and atom's spontaneous emissions}
\label{alg1}
\begin{algorithmic}
%\caption{Pseudocode for register of cavity losses and atom's spontaneous emissions}
\STATE $N$: number of temporal intervals
\STATE $N_A $: number of spontaneous emissions
\STATE $N_C$: number of cavity emissions
\STATE $N_{max}$: maximum number of cavity emissions
\STATE $N_{CA}$: matrix relating the number of cavity emissions with the number of spontaneous emissions

@@

\ENDIF
\ELSE
\STATE $\ket{ \phi(t + \delta t)} = \frac{\ket{\phi^1 (t+ \delta t)}}{\sqrt{1-\delta p}}$
\STATE $t \ += \delta t$
\ENDIF
\ENDFOR
\RETURN  $N_{CA}$
\end{algorithmic}
\end{algorithm}
We compare the collapse probability of Eq.~\eqref{probcol} with a
random number $\epsilon$, uniformly distributed between zero and one.
In the case the condition $\delta p > \epsilon$ is fulfilled we
simulate that a quantum jump has happened, by acting with one of the
collapse operators $a$ or $\sigma_-$ and renormalizing the obtained
state. The operator that will act on the state will be given according
to the relative probabilities of Eqs.~\eqref{relprob}.
\begin{figure}[!t] 
\centering
\includegraphics[scale = 0.5]{distributioneng.pdf}
\caption{Probability distribution of spontaneous emissions for
  different number of cavity photons. Parameters: $\expval{n} \approx 1$, $\kappa = g = 0.1\gamma$. } \label{probdiss}
\end{figure}
If $\delta p > \epsilon$ is not fulfilled we just renormalize the
state, because using a non-hermitian Hamiltonian will have as a
consequence that the probability won't be conserved
\cite{Sakurai:1167961}. For the next step we repeat the procedure
using the new state. One quantum trajectory of total time $t_f$ is
obtained after repeating the previous procedure $t_f/\delta t$ steps.
For each trajectory we count the number of quantum jumps, the jumps
corresponding to operator $a$ reduces the number of photons inside the
cavity and corresponds to a photon that leak the cavity. The jumps
corresponding to operator $\sigma_-$, send the atom back to the ground
state and corresponds to a spontaneous emission. The procedure to
register emissions in the two different channels is summarized in
algorithm. \ref{alg1}.

@@

\begin{center}
\begin{figure}[h!]
\begin{center}
\includegraphics[scale = 0.5]{million2.pdf}
\caption{Pearson correlation coefficient as a function of $\kappa$ for
  quantum trajectories with different duration. Parameters:
  $\expval{n} \approx 1$, $g = 0.1\gamma$.} \label{errorzz}
\end{center}
\end{figure}
\end{center}

@@

\section{Authors contributions}
All the authors were involved in the preparation of the manuscript.
All the authors have read and approved the final manuscript.

\begin{acknowledgement}
This work was supported by DGAPA-UNAM under grant PAPIIT-IG120518.
\end{acknowledgement}

\bibliography{../reftesis}% Produces the bibliography via BibTeX.
